---
title: "Untitled"
format: html
editor: visual
---

# Setting library

```{r}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(janitor)
library(dplyr)
library(GGally)
library(ROCR)
```

# Data tidy

```{r}
datatest <- read.csv("C:\\Users\\10409\\Desktop\\Group29\\dataset29.csv")
datatest <- as.data.frame(lapply(datatest, function(x) gsub(",","",x)))
datatest <- datatest %>%
  filter(across(everything(),~.!="?"))
datatest$Age <- as.numeric(datatest$Age)
datatest$Education <- as.factor(datatest$Education)
datatest$Marital_Status <- as.factor(datatest$Marital_Status)
datatest$Occupation <- as.factor(datatest$Occupation)
datatest$Sex <- as.factor(datatest$Sex)
datatest$Nationality <- as.factor(datatest$Nationality)
datatest$Income <- as.factor(datatest$Income)
datatest$Hours_PW <- as.numeric(datatest$Hours_PW)
```

Observe the correlation of the numeric of data:

```{r}
ggpairs(datatest[,c("Age","Hours_PW","Income")])
```

# Initial impression of the data

## Data levels

```{r}
levels(datatest$Education)
levels(datatest$Marital_Status)
levels(datatest$Occupation)
levels(datatest$Sex)
levels((datatest$Nationality))
```

## Age by Income

```{r}
ggplot(data=datatest,aes(x=Income,y=Age,fill=Income))+
  geom_boxplot()+
  labs(x="Whether larger than 50K",y="Age")+
  theme(legend.position = "none")
```

## Education by Income

```{r}
datatest %>%
  tabyl(Income,Education) %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns()
ggplot(data=datatest,aes(x=Income,group=Education))+
  geom_bar(aes(y=..prop..,fill=Education),stat = "count",position="dodge")+
  labs(x="Whether larger than 50K",y="Proportion")
```

## Martial_Status by Income

```{r}
datatest %>%
  tabyl(Income,Marital_Status) %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns()
ggplot(data=datatest,aes(x=Income,group=Marital_Status))+
  geom_bar(aes(y=..prop..,fill=Marital_Status),stat = "count",position="dodge")+
  labs(x="Whether larger than 50K",y="Proportion")
```

## Occupation by Income

```{r}
datatest %>%
  tabyl(Income,Occupation) %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns()
ggplot(data=datatest,aes(x=Income,group=Occupation))+
  geom_bar(aes(y=..prop..,fill=Occupation),stat = "count",position="dodge")+
  labs(x="Whether larger than 50K",y="Proportion")
```

## Sex by Income

```{r}
datatest %>%
  tabyl(Income,Sex) %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns()
ggplot(data=datatest,aes(x=Income,group=Sex))+
  geom_bar(aes(y=..prop..,fill=Sex),stat = "count",position="dodge")+
  labs(x="Whether larger than 50K",y="Proportion")
```

## Hours_pw by Income

```{r}
ggplot(data=datatest,aes(x=Income,y=Hours_PW,fill=Income))+
  geom_boxplot()+
  labs(x="Whether larger than 50K",y="Hours_PW")+
  theme(legend.position = "none")
```

## Nationality by Income

```{r}
datatest %>%
  tabyl(Income,Nationality) %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns()
ggplot(data=datatest,aes(x=Income,group=Nationality))+
  geom_bar(aes(y=..prop..,fill=Nationality),stat = "count",position="dodge")+
  labs(x="Whether larger than 50K",y="Proportion")
```

# Model selection

```{r}
model_datatest <- glm(Income~Age+Education+Marital_Status+Occupation+Sex+Hours_PW+Nationality,data=datatest,family = binomial(link="logit"))
model_datatest %>% summary()
```

```{r}
model_datatest2 <- glm(Income~Age+Education+Marital_Status+Occupation+Hours_PW,data=datatest,family = binomial(link="logit"))
model_datatest2 %>% summary()
```

Check the fitted of the model:

```{r}
qchisq(p=0.95,df=1351)
```

We can find it fitted well, the residual deviance of model2 is 998.6 which is lower than 1437.623.

### Option choose of model

```{r}
data.new <- datatest

data.new$Education <- as.factor(ifelse(data.new$Education %in% c("1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th", "12th"), "Low_Education", as.character(data.new$Education)))
data.new$Education <- as.factor(ifelse(data.new$Education %in% c("Assoc-acdm","Assoc-voc","Bachelors","Doctorate","HS-grad","Masters","Prof-school","Some-college"), "High_Education", as.character(data.new$Education)))

data.new$Marital_Status <- as.factor(ifelse(data.new$Marital_Status %in% c( "Separated", "Widowed","Never-married","Divorced"), "Other", as.character(data.new$Marital_Status)))
data.new$Marital_Status <- as.factor(ifelse(data.new$Marital_Status %in% c("Married-spouse-absent","Married-civ-spouse"), "Married", as.character(data.new$Marital_Status)))


data.new$Occupation <- as.factor(ifelse(data.new$Occupation %in% c("Exec-managerial","Prof-specialty", "Tech-support"), "High-Income Jobs", as.character(data.new$Occupation)))
data.new$Occupation <- as.factor(ifelse(data.new$Occupation %in% c("Craft-repair", "Sales", "Transport-moving", "Machine-op-inspct"), "Middle-Income Jobs", as.character(data.new$Occupation)))
data.new$Occupation <- as.factor(ifelse(data.new$Occupation %in% c("Adm-clerical","Protective-serv","Handlers-cleaners","Farming-fishing","Other-service", "Priv-house-serv"), "Low-Income Jobs", as.character(data.new$Occupation)))

model_new <- glm(Income ~ Age + Education + Marital_Status +Occupation+ Hours_PW,data = data.new, family = binomial(link="logit"))
summary(model_new)

```

```{r}
qchisq(p=0.95,df=1375)
```

## Odds of model

```{r}
plot_model(model_datatest2, show.values = TRUE,
           title = "", show.p = FALSE, value.offset = 0.5)
```

```{r}
plot_model(model_new, show.values = TRUE,
           title = "", show.p = FALSE, value.offset = 0.5)
```

## ROC curves

```{r}
predicted <- predict(model_datatest2,type = "response")
predicted_class <- ifelse(predicted>0.5,">50k","<=50k")
score <- prediction(predicted,datatest$Income)
perf <- performance(score,"tpr","fpr")
auc <- performance(score,"auc")
perfd <- data.frame(x=perf@x.values[1][[1]],y=perf@y.values[1][[1]])
ggplot(perfd,aes(x=x,y=y))+geom_line()+
  xlab("False positive rate") + ylab("True positive rate") +
  ggtitle(paste("Area under the curve:", round(auc@y.values[[1]], 3)))
```

Area under the curve is 0.882.

```{r}
predicted2 <- predict(model_new,type = "response")
predicted_class2 <- ifelse(predicted2>0.5,">50k","<=50k")
score2 <- prediction(predicted2,data.new$Income)
perf2 <- performance(score2,"tpr","fpr")
auc2 <- performance(score2,"auc")
perfd2 <- data.frame(x=perf2@x.values[1][[1]],y=perf2@y.values[1][[1]])
ggplot(perfd2,aes(x=x,y=y))+geom_line()+
  xlab("False positive rate") + ylab("True positive rate") +
  ggtitle(paste("Area under the curve:", round(auc2@y.values[[1]], 3)))
```

Area under the curve is 0.853.
